The `Maltreatment and Abuse Chronology of Exposure' (MACE) scale is essentially unidimensional

### 2022-07-28 manuscript/main.tex
updates to manuscript

The MACE scale contains 12 sets of two to three items with response dependencies (Table S1). Some item sets share a structural dependency, where a response to one item logically requires or strongly implies a response on a second. For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes''). Other item sets are characterized by redundancy dependency, where the same question is essentially asked twice using slightly different language. One such example is item 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``).

For the current analyses, we were concerned about response dependence for two reasons. First, response dependence can inflate estimates of item discrimination parameters, thereby making items appear more informative than they actually are \citep{marais2008formalizing}. Second, in the context of bifactor models, specific factors may become contaminated by dependent items thereby making them appear more reliable than they actually are \citep{reise2013applying}. Following \cite{marais2008formalizing}, we therefore converted each set of dependent dichotomous items into a single polytomous item, thereby reducing the original 52-item MACE scale to 39 items.

@@

In order to evaluate the dimensionality of the MACE scale, we fit a series of four confirmatory item factor models to the endorsement data of each sample separately. The basis of each model was the The basis of each model was the graded response model for ordered polytomous data \citep{samejima1997graded}. The structure of each model is depicted in Figure \ref{fig:models}. The first model is a one-factor (undimensional) model whereby all items load onto a general maltreatment factor (Figure \ref{fig:models}a). The second is a bifactor model with a general maltreatment factor and 10 specific factors corresponding to each subscale of the MACE (Figure \ref{fig:models}b).

The third model --- inspired by the deprivation-threat dimensional model \citep{mclaughlin2014childhood} --- is a bifactor S-1 model with a general maltreatment factor and one specific factor measuring neglect (Figure \ref{fig:models}c). We first attempted to fit this structure using a conventional bifactor model with two specific factors (i.e. threat, deprivation), but we observed vanishing factor loadings on the threat specific factor. Therefore, we instead utilized the bifactor S-1 model structure \citep{eid2017anomalous} where all of the threat (abuse) items comprise the general or reference factor, and the neglect items load onto both the general and specific factors.

The fourth model attempts to provide a more parsimonious explanation of the MACE scale response data. We hypothesized items on the MACE can be summarized by overarching factors: maltreatment by adults in the home and maltreatment by peers. We also propose including a methods factor to account for unexplained variance in endorsements specific to the reverse-scored items. Thus, we tested a bifactor S-1 model with a general maltreatment factor and two specific factors (i.e. peer victimization, reverse-scoring; Figure \ref{fig:models}d). Similar to Model 3, initial fits of this structure using a conventional bifactor model with three specific factors resulted in vanishing factor loadings on the household abuse specific factor.

All the confirmatory item factor models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.26; \citealt{carpenter2017stan}). For all models, four separate chains with randomized start values each took 3,000 samples from the posterior. The first 2,000 samples from each chain were discarded, so that 4,000 post-warmup samples from the posterior were retained. The $\hat{R}$ values for all parameters were less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.

@@

We relied on several indices to evaluate the fit of the models to the data. First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited, maydeu2013goodness}, which is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the number of reduced first- and second-order marginal residuals (i.e. $\kappa = n(n + 1)/2$ for $n$ items) and $\nu$ denotes the number of free parameters. (To improve the stability of the $M_2$, the polytomous items were binarized, corresponding to no versus any endorsement of maltreatment.)  Based on the $M_2$ statistic, we computed the root mean square error of approximation ($\text{RMSEA}_2$), comparative fit index (CFI), and Tucker-Lewis index (TLI). Following convention \citep{hu1999cutoff}, relatively good fit is indicated by the following indices: RMSEA < 0.06, CFI > 0.95, and TLI > 0.95.

Given the questionable utility of SEM-based fit indices for item response theory applications (e.g. \citealt{reise2014evaluating}), we calculated two additional fit indices. First, we relied on posterior predictive model checking and a $\chi^2_{NC}$ discrepancy measure based on the total score distribution \citep{sinharay2006posterior}. This index compares the observed and expected proportion of participants who endorse maltreatment at each sum-score level. To detect local dependence, we calculated Yen's $Q_3$ statistic for each pair of items \citep{yen1984effects}. Following \cite{christensen2017critical}, we calculated the critical $Q_3$-value (i.e. the value above which local dependence is indicated) for each model and sample. Specifically, for each model and sample, we simulated 1000 locally-independent response datasets using the posterior distribution of item and ability parameters. We then recorded maximum $Q_3$ value per simulated dataset. The critical $Q_3$-value (per sample and dataset) was defined as the 99th percentile of the max $Q_3$ values.

@@

\subsection{Reliability indices}

>>>>

The MACE contains a total of 12 item pairs and triplets characterized by response dependence (Table S1). The items in some of these sets are structurally dependent, where a response to one item requires, or strongly implies, a certain response on the second. For example, a person endorsing item 8 (``Parents or guardians hit you so hard that you received medical attention'') is all but certain to endorse item 7 (``Parents or guardians hit you so hard that it left marks for more than a few minutes''). Other items are dependent by virtue of redundancy. One such example is item 36 (``Peers forced you to engage in sexual activity against your will'') and 37 (``Peers forced you to do things sexually that you did not want to do``); here the same question is asked twice using slightly different language.

For this study, we were concerned about response dependence for two reasons. First, it can inflate the estimates of item discrimination parameters, making items appear more informative than they actually are \citep{marais2008formalizing}. Second, response dependence can inflate the estimates of item loadings on specific factors in bifactor models, giving false support for a larger number of factors than necessary \citep{reise2013applying}. Following \cite{marais2008formalizing}, we therefore converted each set of dependent items into a single polytomous item. Thus, the MACE was reduced from 52 to 39 items in total.

@@

To begin our investigation of the MACE factor structure, we fit a series of confirmatory item factor models to the response data from each sample separately. The basis of each model is the graded response model for ordered polytomous data \citep{samejima1997graded}. The factor structure for each of the four models is depicted in Figure \ref{fig:models}. The first model we fit is a one-factor (undimensional) model whereby all items load onto a general maltreatment factor (Figure \ref{fig:models}a). The second is a bifactor model, in which all items load onto a general maltreatment factor and also one of 10 specific factors corresponding to their subscale membership (Figure \ref{fig:models}b).

The third model we fit investigated a conceptual model of maltreatment that distinguishes between threat and deprivation as distinct forms of childhood adversity \citep{mclaughlin2014childhood}. Initially we fit a bifactor model in which all items loaded onto a general maltreatment factor one either the threat or deprivation specific factors. However, we observed vanishing item loadings on the threat specific factor. As such, we instead fit a bifactor S-1 model \citep{eid2017anomalous} where all items load onto a general maltreatment (or reference) factor, and only the items measuring neglect experiences load onto their own specific factor (Figure \ref{fig:models}c).

The fourth and final model we fit was designed provide a more parsimonious factor structure for the MACE. Items in the MACE can be neatly divided according to the source of adversity: maltreatment by adults at home (e.g. parents) or by peers. We hypothesized then that there would be two principal factors underlying responses on the MACE, maltreatment by adults and peer victimization. Based on preliminary model fits, we also suspected the presence of a secondary methods factor affecting the 6 reverse-scored items. Initially we fit a bifactor model in which all items loaded onto a general maltreatment factor and also one of three specific factors: maltreatment by adults, peer victimization, or reverse-scoring. However, we also observed vanishing loadings for the maltreatment by adults specific factor. Thus, we instead fit a bifactor S-1 model where all items load onto a general maltreatment factor; in addition, the peer victimization and reverse-scoring items load onto their own respective specific factors (Figure \ref{fig:models}d).

All confirmatory item factor models were estimated within a Bayesian framework using Hamiltonian Monte Carlo as implemented in Stan (v2.26; \citealt{carpenter2017stan}). For each model, four separate chains with randomized start values each took 3,000 samples from the posterior. The first 2,000 samples from each chain were discarded, so that 4,000 post-warmup samples from the posterior were retained. The $\hat{R}$ values for all parameters were less than 1.01, indicating acceptable convergence between chains, and there were no divergent transitions in any chain.

@@

We relied on multiple indices to evaluate the fit of the models to the data. First we calculated three traditional goodness-of-fit statistics using the ordinal $M_2$ statistic \citep{cai2013limited, maydeu2013goodness}. $M_2$ is asymptotically $\chi^2$-distributed with $\kappa − \nu$ degrees of freedom ($df$), where $\kappa$ is the reduced number of first- and second-order marginal residuals (i.e. $\kappa = n(n + 1)/2$ for $n$ items) and $\nu$ denotes the number of free parameters. To improve the stability of the $M_2$ statistic, the responses to polytomous items were binarized (i.e. none vs. some maltreatment). Using the $M_2$ statistic, we calculated the root mean square error of approximation (RMSEA), comparative fit index (CFI), and Tucker-Lewis index (TLI) for each model fit. Following convention \citep{hu1999cutoff}, the benchmarks for adequate model fit are RMSEA < 0.08, CFI > 0.90, and TLI > 0.90; in turn, the benchmarks for good model fit are RMSEA < 0.05, CFI > 0.95, and TLI > 0.95.

Given the questionable utility of SEM-based fit indices for item response models (e.g. \citealt{reise2014evaluating}), we calculated two additional fit indices. First, we employed posterior predictive model checking and the $\chi^2_{NC}$ discrepancy measure based on the overall severity score distribution \citep{sinharay2006posterior}. This measure compares the observed and model-predicted proportion of participants at each overall severity score level. Second, we calculated Yen's $Q_3$ statistic for each pair of items to detect local dependence \citep{yen1984effects}. For each model and sample, we also calculated the critical $Q_3$-value above which local dependence is indicated  \citep{christensen2017critical}. Briefly, for each model and sample, we simulated 1000 locally independent datasets using the posterior distribution of the model parameters. We then recorded the maximum $Q_3$ value per simulation. The critical $Q_3$-value was defined as the 99th percentile of these max $Q_3$ values.

@@

\subsection{Model-based reliability indices}
